---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load FAS Litters Data

1. Load csv data
2. Clean variables
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

# DPLYR SELECT

## Dplyr (Select): choose columns of interest 

1. For select, first specify dataframe
2. Then specify columns you'd like to keep. In this example, we want to keep "group" and "litter_number" column.
```{r}
select(litters_df, group, litter_number)
```

1. select then specify dataframe (litters_df)
2. specify columns to keep (keep columns from gd0_weight to gd_of_birth)
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

## Dplyr (Select): choose columns you'd like to remove

1. specify dataframe (litters_df)
2. indicate column you'd like to remove by placing "-" before column name (ex: -litter_number = remove litter number column)

```{r}
select(litters_df, -litter_number)
```

## Renaming columns

* OPTION 1: Use select columns and rename selected columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

* OPTION 2: Use rename to only rename columns and not select

```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

## Select variable of interest that starts or ends with certain term

* select everything that starts with gd
```{r}
select(litters_df, starts_with("gd"))
```

## Move variable of interest to the front

* OPTION 1: Using SELECT
move column of interest to the front (i.e., make litter_number be the first column)

```{r}
select(litters_df, litter_number, everything())
```

* OPTION 2: Using RELOCATE
make litter_number come to the front

```{r}
relocate(litters_df, litter_number)
```

# DPLYR FILTER --> requires conditions

## Greater than, GE to, equal to

1. example 1: only select subjects whose gd0_weight is less than 22
```{r}
filter(litters_df, gd0_weight < 22)
```

2. example 2: only select those whose gd0_weight is ge to 22
```{r}
filter(litters_df, gd0_weight >= 22)
```

3. example 3: select those whose gd_of_birth is equal to 20
```{r}
filter(litters_df, gd_of_birth == 20)
```

## Negation statements

1. gd_of_birth not equal to 20
```{r}
filter(litters_df, !(gd_of_birth == 20))
```

2. two conditions using and
```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth ==20)
```

3. filtering charater variables
```{r}
filter(litters_df, group == "Mod8")
```

4. filtering using "this or that"
Below example would mean "select those in group that are either Con7 or Mod8.
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

# DPLYR MUTATE --> modify existing var or create new var

## Creating new variables, modifying variables
1. create a new variable: wt_gain
2. change group variable: change everything in group variable to lowercase

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```

# DPLYR ARRANGE --> putting things in order

1. example 1: arrange by pups_born_alive in ascending order
```{r}
arrange(litters_df, pups_born_alive)
```

2. example 2: arrange 2 variables
```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```

