---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Read in some data (IMPORTING CSV FILE)

Read in the litters dataset (IMPORTING CSV FILE)
1. litters_df = imported CSV will become a data frame that is named "litters_df"
2. read_csv = pacakge w/in readr
3. "./data/FAS_litters.csv" = pathway
4. overwrite litters_df by cleaning the names using "clean_names(litters_df)"
5. janitors::clean_names --> we are using clean_names within janitor package
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## Take a look at the data

Option 1:
```{r}
litters_df
```

Option 2: head only gives first 6 rows
```{r}
head(litters_df)
```

Option 3: tail only gives last 6 rows
```{r}
tail(litters_df)
```

Option 4: skim --> gives mean, sd, etc.
```{r}
skimr::skim(litters_df)
```

Option 5: View(litters_df) within the console

## Options to read_csv

1. skip = 10 --> skip the first 10 rows
```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
```

2. specifying missing values: blank, NA, 999, . = missing values
```{r}
litters_df = read_csv("./data/FAS_litters.csv", na = c("","NA", 999,"."))
```

## IMPORTING (READING IN) EXCEL FILE

```{r}
library(readxl)
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
```

1. specifying range of cells to be read in
```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

## IMPORTING SAS FILE

```{r}
library(haven)
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
pulse_df = janitor::clean_names(pulse_df)
pulse_df
```

## Comparison with Base R

1. Always use read_csv NOT read.csv. read.csv will read in additional unnecessary info (i.e., space, comma)
