---
title: "Tidy Data"
output: github_document
---

```{r setup}
library(tidyverse)
```

## 'povot_longer'

Load the PULSE data

```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Wide format to long format...

1. pivot_longer: change wide format to long format. Here, we want to combine bdi columns into one row
2. names_to: place bdi variables to "visit" column
3. names_prefix: remove prefix "bdi_score_"
4. values_to: move bdi values to "bdi" column

```{r}
pulse_data_tidy =
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )
```

rewrite, combined, and extend (to add a mutate)

1. use relocate so id & visit columns appear first
2. use mutate to change visit values to numbers via `recode`
```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
) %>% 
    relocate(id, visit) %>% 
    mutate(visit = recode(visit, "bl" = "00m"))
```


