---
title: "Tidy Data"
output: github_document
---

```{r setup}
library(tidyverse)
```

## 'pivot_longer'

Load the PULSE data

```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Wide format to long format...

1. pivot_longer: change wide format to long format. Here, we want to combine bdi columns into one row
2. names_to: place bdi variables to "visit" column
3. names_prefix: remove prefix "bdi_score_"
4. values_to: move bdi values to "bdi" column

```{r}
pulse_data_tidy =
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )
```

rewrite, combined, and extend (to add a mutate)

1. use relocate so id & visit columns appear first
2. use mutate to change visit values to numbers via `recode`
```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
) %>% 
    relocate(id, visit) %>% 
    mutate(visit = recode(visit, "bl" = "00m"))
```

## 'pivot wider'

Make up some data!

1. make data using tibble
2. using pivot wider:
    a. get column names from "time"
    b. get values for time from "mean"

```{r}
analysis_result =
  tibble(
    group = c("tx", "tx", "p", "p"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )

analysis_result %>% 
  pivot_wider(
    names_from = "time",
    values_from = "mean"
)
```

## stacking tables together (aka merging tables)

### binding rows

Using LotR data.
1. import each table

```{r}
fellowship_ring =
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers =
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king =
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

2. bind all the rows together 

```{r}
lotr_tidy = 
    bind_rows(fellowship_ring, two_towers, return_king) %>% 
    janitor::clean_names() %>% 
    relocate(movie) %>% 
    pivot_longer(
      female:male,
      names_to = "gender",
      values_to = "words"
    )
```

## join dataset

### import & clean

Pups_df
1. Import the FAS datasets.
2. Data cleaning

Litters_df
1. import
2. data clean
3. split treatment period & control in group variable using `separate`

```{r}
pups_df =
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  relocate(litter_number) %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)
```

### join datasets

```{r}
fas_df =
  left_join(pups_df, litters_df, by = "litter_number") %>% 
  arrange(litter_number) %>% 
  relocate(litter_number, dose, day_of_tx)
```

